---
output:
  pdf_document: default
  html_document: default
---
---
title: "Lesson 8 Practical Exercise"
author: "Austin Logan"
footer: `r paste('Lesson 8 Practical Exercise')`
date: "`r format(as.Date(c('2021/04/21')), '%d %B %Y')`"
output:
  pdf_document:
    df_print: kable
  html_document:
    df_print: paged
  slidy_presentation:
    code_folding: hide
    fig_caption: yes
    smart: no
keep_tex: yes
graphics: yes
---

# 1. Preliminaries

```{r intro, echo=FALSE,message=FALSE, warning=FALSE}
options(repos="https://cran.rstudio.com" )
library(xlsx)
library(printr)
library(xtable)
library(formatR)
library(nlme)
options(xtable.comment = FALSE, xtable.type="latex")
```

# Perform data housekeeping - upload data for Example 5.6, Example 5.7, and Printing Ink; name columns, display to make sure it reads properly, etc. 

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
file=file.path(getwd(),"data-ex-PrintingInk.xlsx")
printData=read.xlsx(file,1)
names(printData)=c("x1","x2","x3","y_bar","sdev")
attach(printData)
printData
```

# 2. Example 5.5 Method (using Printing Ink data) 

## a) Create a linear model relating the regressors to the standard deviation ('s' in th data file)

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
inkModel=lm(sdev~x1+x2+x3)
summary(inkModel)
```

## b) Define model weights

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
printData$weights=1/(inkModel$fitted.values^2+1)
```

## c) Solve weighted least squares using matrices

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
anv=anova(inkModel)
V=anv$`Mean Sq`*diag(1/printData$weights)
W=solve(V)
ones=matrix(1,27,1)
X=matrix(c(ones,x1,x2,x3),27,4)
beta=solve(t(X)%*%W%*%X)%*%t(X)%*%W%*%y_bar
beta
```

## d) Solve weighted least squares using lm() and entering weights


```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
weightModel=lm(y_bar~x1+x2+x3,weights=printData$weights)
summary(weightModel)
```

## e) Display both sets of coefficients - should be the same (minor differences possible due to rounding)

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}
beta
weightModel$coefficients
```

# 3. Example 5.6 (p. 196-199)

## a) Create a linear model predicting time as a function of aspect, paper,and the interaction. Results should match Table 5.11.

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

```

## b) Create a linear model of groups, with the response being the group average (i.e. use Table 5.14 as input data). Results should match Table 5.15.

```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

```

# 4. Example 5.7 (p.201-202)

## a) Run regression model using City as a random factor using lme(). Compare StdDev for (Intercept) and Residual to variances in Figure 5.12.


```{r, echo=T, tidy=TRUE, tidy.opts=list(width.cutoff=60)}

```
